Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2015-12-30 reviews/response-jarosch.tex

\begin{itemize}
    \item{Our \emph{default} configuration used in the control run and all other
          simulations in the manuscript include rheologic parameters, $A_c$,
          $A_w$, $Q_c$ and $Q_w$, derived from \citet{Paterson.Budd.1982} and
          given in \citet[Eqn.~5]{Bueler.Brown.2009}, and $E_{SIA}=1$.}
    \item{Our \emph{hard ice} configuration include rheologic parameters,
          $A_c$, $A_w$, $Q_c$ and $Q_w$, derived from
          \citet[p.~72 and 76]{Cuffey.Paterson.2010}, and $E_{SIA}=1$, which
          correspond to a stiffer rheology than that used in the control run.}
    \item{Our \emph{soft ice} configuration include rheologic parameters,
          $A_c$, $A_w$, $Q_c$ and $Q_w$, derived from
          \citet[p.~72 and 76]{Cuffey.Paterson.2010}, and $E_{SIA}=5$, the
          recommended value for ice age polar ice
          \citep[p.~77]{Cuffey.Paterson.2010}.}
\end{itemize}

An additional simulation with rheologic parameters, $A_c$, $A_w$, $Q_c$ and
$Q_w$, derived from \citet{Cuffey.Paterson.2010}, and $E_{SIA}=2$, the
recommended value for Holocene polar ice \citep[p.~77]{Cuffey.Paterson.2010}
was performed, but its results were very similar to that of our default run,
thus we decided to not present it here.

Actual parameter values for $A_c$, $A_w$, $Q_c$, $Q_w$ and $E_{SIA}$ used in
our simulations are given in Table~\ref{tab:sens_params}, while the effect of
the three different parametrisations on temperature-dependent ice softness,
$A$, is illustrated in Fig.~\ref{fig:sens_plot_rheo}.

\begin{table*}
  \caption{Parameter values used in the sensitivity test.}
  \label{tab:sens_params}
  \centering\makebox[\textwidth]
  {\begin{tabular}{l|ccccc|cc|cc}
    \tophline
            & \multicolumn{5}{c|}{Rheology}
            & \multicolumn{2}{c|}{Sliding}
            & \multicolumn{2}{c}{GRIP scaling} \\
    Config. & $A_c$ & $A_w$ & $Q_c$ & $Q_w$ & $E_{SIA}$
            & $\delta$ & $W_{max}$ & $f$ & $T_{[32, 22]}$ \\
            & \multicolumn{2}{c}{(\unit{Pa^{-3}\,s^{-1}})}
            & \multicolumn{2}{c}{(\unit{J\,mol^{-1}})}
            & & & (m) \\
    \middlehline
    Default$^1$  & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.02 & 2 & 0.38 & 6.2 \\
    \middlehline
    Soft ice$^2$ & $2.847\times 10^{-13}$
                 & $2.356\times 10^{-2}$
                 & $   60\times 10^3$
                 & $  115\times 10^3$
                 & 5 & 0.02 & 2 & 0.40 & 6.6 \\
    Hard ice$^2$ & $2.847\times 10^{-13}$
                 & $2.356\times 10^{-2}$
                 & $   60\times 10^3$
                 & $  115\times 10^3$
                 & 1 & 0.02 & 2 & 0.37 & 6.0 \\
    \middlehline
    Soft bed     & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.01 & 1 & 0.40 & 6.5 \\
    Hard bed     & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.05 & 5 & 0.36 & 5.9 \\
    \bottomhline
  \end{tabular}}
  \belowtable{After $^1$\citet{Paterson.Budd.1982,Bueler.Pelt.2015};
              and $^2$\citet{Cuffey.Paterson.2010}.}
\end{table*}

@@

\clearpage  % moving on to basal sliding

In our simulations, basal sliding is governed by a~pseudo-plastic sliding law,
already given in the manuscript but recalled here for the sake of completeness,
%
\begin{equation}
    \label{eqn:pseudoplastic}
    \vec{\tau}_b = -\tau_c \frac{\vec{v}_b}{{v_{th}}^q\,|\vec{v}_b|^{1-q}} \,,
\end{equation}
%
which relates the bed-parallel shear stresses, $\vec{\tau}_b$, to the sliding
velocity, $\vec{v}_b$. The yield stress, $\tau_c$, is modelled using the
Mohr--Coulomb criterion,
%
\begin{equation}
   \tau_c = c_0 + N\,\tan{\phi} \,,
\end{equation}
%
where cohesion, $c_0$, is assumed to be zero. Effective pressure, $N$,
is related to the ice overburden stress, $\rho gh$, and the modelled amount of
subglacial water, using a formula derived from laboratory experiments with till
extracted from the base of Ice Stream B in West Antarctica
\citep[Eqn.~23]{Tulaczyk.etal.2000, Bueler.Pelt.2015},
\begin{equation}
    N = \delta \rho gh \, 10^{(e_0/C_c) (1 - (W/W_{max}))} \,,
\end{equation}
where $\delta$ sets the minimum ratio between the effective pressure, $N$, and
the ice overburden stress, $\rho gh$, $e_0$ is a measured reference void ratio
and $C_c$ is a measured compressibility coefficient. The amount of water at the
base, $W$, varies from zero to $W_{max}$, a threshold above which additional
melt water is assumed to drain off instantaneously.

In our sensitivity test, we set constant the pseudo-plastic sliding exponent,
$q=0.25$, and the threshold velocity, $v_{th}=100$\,\unit{m\,s^{-1}}, according
to values used by \citet{Aschwanden.etal.2013}, the till cohesion, $c_0=0$,
whose measured values are consistently negligible
\citep[p.~268]{Tulaczyk.etal.2000, Cuffey.Paterson.2010}, the till reference
void ratio, $e_0=0.69$, and the till compressibility coefficient, $C_c=0.12$,
according to the only measurements available to our knowledge by
\citep{Tulaczyk.etal.2000}.

We also use a constant spatial distribution of the till friction angle, $\phi$,
whose values vary from 15 to 45{\degree} as a piecewise-linear function of
modern bed elevation, with the lowest value occurring below modern sea level
(0\,m above sea level, m~a.s.l.) and the highest value occurring above the
generalised elevation of the highest shorelines
\citep[200\,m~a.s.l.,][Fig.~5]{Clague.1981}. This range of value span over the
range of measured values for glacial till of 18 to 40{\degree}
\citep[p.~268]{Cuffey.Paterson.2010}. It accounts for frictional basal
conditions associated with discontinuous till cover at high elevations, and
a~weakening of till associated with the presence of marine sediments at low
elevations (cf. \citealp{Martin.etal.2011};
\citealp[supplement]{Aschwanden.etal.2013}; \citealp{PISM-authors.2015}).

@@

On the other hand, we test different values for the minimum ratio between the
effective pressure and the overburden stress, $\delta$, and the maximum water
height in the till, $W_{max}$, as follow.

\begin{itemize}
    \item{Our \emph{default} configuration used in the control run and all other
          simulations in the manuscript include $\delta=0.02$ and
          $W_{max}=2$\,m as in \citet{Bueler.Pelt.2015}.}
    \item{Our \emph{soft bed} configuration use $\delta=0.01$ and
          $W_{max}=1$\,m.}
    \item{Our \emph{hard bed} configuration use $\delta=0.05$ and
          $W_{max}=5$\,m.}
\end{itemize}

The effect of the three different parametrisations on the effective pressure on
the till, $N$, in response to water content, $W$, is illustrated in
Fig.~\ref{fig:sens_plot_ntil}.
All parameter choices are recalled in Table~\ref{tab:sens_params}.

Finally, we adjusted the GRIP linear scaling factor for each run, so that they
result in a similar glaciated area at the Last Glacial Maximum (LGM) to that
modelled with the default configuration (Table~\ref{tab:sens_params}).
In other words, the sensitivity in modelled ice sheet extent at MIS~2 is
expressed through the scaling factor required to obtain a fixed target area,
since we consider this scaling factor as a free parameter in our study.

@@

\begin{figure}
    \centering
    \includegraphics{sens_plot_ntil}
    \caption{Effective pressure, $N$, as a function of water height at the
             base, $W$, for the default ($\delta=0.02$, $W_{max}=2$\,m),
             hard bed ($\delta=0.05$, $W_{max}=5$\,m), and soft bed
             ($\delta=0.01$, $W_{max}=1$\,m) sliding parametrisations, using a
             linear scale (top panel) and a logarithmic scale (bottom panel).}
    \label{fig:sens_plot_ntil}
\end{figure}

\clearpage  % moving on to the results

By analogy to our manuscript's Table~3 (ice volume and extent extrema) and
Fig.~3 (sea-level relevant ice volume time series), the results of our
sensitivity study are presented here using a similar layout in
Table~\ref{tab:sens_extrema} and Fig.~\ref{fig:sens_ts}.

As a result of the different scaling factors applied, the resulting simulations
show very little difference in the modelled glaciated areas corresponding to
maximum ice volumes during MIS~2, but also during MIS~4
(Table~\ref{tab:sens_extrema}). However, this can not be said for the modelled
glaciated area corresponding to minimum ice volume during MIS~3. In fact, the
extent of the remnant ice cap which persists over the Skeena Mountains during
this stage shows a significant sensitivity to ice rheology of 31\%, and
an even more important sensitivity to basal sliding parameters of 62\%
(Table~\ref{tab:sens_extrema}).

The modelled sea-level relevant ice volumes show more variability than the
modelled glaciated areas (Table~\ref{tab:sens_extrema},
Fig.~\ref{fig:sens_ts}). As could be expected, softer ice and weaker till both
result in a thinner ice sheet, while harder ice and stronger till result in a
thicker ice sheet. For instance, peak ice volume during the MIS~2 (LGM) varies
by 30\% between the two parametrisations of ice rheology used, and by 21\%
between the two parametrisations of till properties used. The differences in
sea-level relevant ice volume are greatest during the MIS~3
(Table~\ref{tab:sens_extrema}, Fig.~\ref{fig:sens_ts}) where both the areal and
thickness contributions add up.

\begin{table*}
  \caption{Extremes in Cordilleran ice sheet volume and extent corresponding to
           MIS~4, 3 and 2 using the GRIP paleo-climate forcing with each
           parameter configuration (Fig.~\ref{fig:sens_ts}). Relative
           differences (r. diff.) give rough error estimates related to
           varying selected ice rheology and basal sliding parameters
           (Table.~\ref{tab:sens_params}).}
  \label{tab:sens_extrema}
  \centering\makebox[\textwidth]
  {\begin{tabular}{l*{3}{|ccc}}
    \tophline
             & \multicolumn{3}{c}{Age (ka)}
             & \multicolumn{3}{c}{Ice extent (\unit{10^6\,km^2})}
             & \multicolumn{3}{c}{Ice volume (m~s.l.e.)} \\
    Config.  &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2
             &  MIS~4 &  MIS~3 &  MIS~2 \\
    \middlehline
    Default  &  57.59 &  42.91 &  19.14
             &   1.93 &   0.67 &   2.09
             &   7.43 &   1.54 &   8.62 \\
    \middlehline
    Soft ice &  58.89 &  49.97 &  21.57
             &   1.96 &   0.54 &   2.08
             &   6.58 &   1.03 &   6.88 \\
    Hard ice &  57.32 &  42.90 &  19.14
             &   1.90 &   0.75 &   2.12
             &   7.83 &   1.91 &   9.46 \\
    R. diff. &    3\% &   16\% &   13\%
             &    3\% &   31\% &    2\%
             &   17\% &   57\% &   30\% \\
    \middlehline
    Soft bed &  58.90 &  49.21 &  19.53
             &   1.88 &   0.55 &   2.05
             &   6.46 &   1.03 &   7.52 \\
    Hard bed &  57.31 &  42.91 &  19.14
             &   1.93 &   0.96 &   2.13
             &   7.99 &   2.89 &   9.31 \\
    R. diff. &    3\% &   15\% &    2\%
             &    3\% &   62\% &    4\%
             &   21\% &  120\% &   21\% \\
    \bottomhline
  \end{tabular}}
  \belowtable{}
\end{table*}

\begin{figure*}
  \centering
  \includegraphics{sens_ts}
  \caption{Modelled sea-level relevant ice volume through the last 120\,ka
           in the simulation forced by the GRIP paleo-climate record, using
           default parameters (black curves), different ice rheology parameters,
           (top panel) and different basal sliding parameters (bottom panel).
           Gray fields indicate Marine Oxygen Isotope Stage (MIS) boundaries
           for MIS~2 and MIS~4 according to a global compilation of benthic
           \chem{\delta^{18}O} records \citep{Lisiecki.Raymo.2005}.}
  \label{fig:sens_ts}
\end{figure*}

@@

% ----------------------------------------------------------------------

\sechead{2 \quad Specific Comments}

\referee{%
    I refer to text locations in the discussion paper by page number (P) and
    line number range (L) for the specific comments.}

\referee{%
    \textbf{P4151 L11-16:} In this sentence the authors refer back to their
    previous work \citep{Seguinot.etal.2014} and highlight that the NARR
    temperature and precipitation fields are the most suitable present day
    climate datasets to be used. Especially since the NARR precipitation fields
    include steep precipitation gradients which are required as identified by
    et \citet{Seguinot.etal.2014}. NARR is delivered on a 32\,km Lambert grid,
    and thus it is questionable how ``steep'' these gradients can be, given
    the rather smooth representation of the existing topography on a 32\,km
    grid. \citet{Seguinot.etal.2014} have partly discussed that however. NARR
    precipitation and temperature fields have been evaluated in detail based on
    available station data for large parts of the study domain dealt with in
    this manuscript. This evaluation \citep{Jarosch.etal.2010} demonstrated
    that NARR has difficulties simulating orographic processes in the Coast
    Mountains which in turn results in unrealistic atmospheric conditions over
    the Rocky Mountains. \citet{Jarosch.etal.2010} further concluded that
    physics based downscaling is required to adequately drive glacier models in
    that region. The authors should argue in more detail here why they think
    that NARR precipitation fields at 32\,km are adequate to drive their model
    and reflect their arguments with the findings of \citet{Jarosch.etal.2010}.
    A solid argument here is of special importance as the authors assume the
    present day precipitation fields to be valid throughout their model time
    period (120\,ky to present) without further corrections (cf. section~2.4
    equation~6).}

### 2015-12-30 reviews/response-jarosch.tex

\begin{figure*}
  \centering
  \includegraphics{sens_ts}
  \caption{Modelled sea-level relevant ice volume through the last 120\,ka
           in the simulation forced by the GRIP paleo-climate record, using
           default parameters (black curves), different ice rheology parameters,
           (top panel) and different basal sliding parameters (bottom panel).
           Gray fields indicate Marine Oxygen Isotope Stage (MIS) boundaries
           for MIS~2 and MIS~4 according to a global compilation of benthic
           \chem{\delta^{18}O} records \citep{Lisiecki.Raymo.2005}.}
  \label{fig:sens_ts}
\end{figure*}

\clearpage  % eventually moving on to the next point

All the information detailed above, including
Eqns~\ref{eqn:glenslaw} and~\ref{eqn:softness}, default parameter values,
Tables~\ref{tab:sens_params} and~\ref{tab:sens_extrema}, and
Figs.~\ref{fig:sens_plot_rheo}--\ref{fig:sens_ts} will be included in Sect.~2
(Model setup) or in the new section (Sensitivity to ice flow parameters).
Relevant discussion points in Sect.~4(Comparison with the geologic record)
will be updated to account for these new results.

### 2015-12-30 reviews/response-jarosch.tex

\begin{figure}
    \centering
    \includegraphics{sens_plot_rheo}
    \caption{Ice softness parameter, $A$, as a function of pressure-adjusted
             temperature, $T_{pa}$, for the default \citet{Paterson.Budd.1982},
             hard ice \citet{Cuffey.Paterson.2010}, and soft ice
             \citet[with $E_{SIA}=5$]{Cuffey.Paterson.2010} rheologies, using
             a linear scale (top panel) and logarithmic scale (bottom panel).}
    \label{fig:sens_plot_rheo}
\end{figure}

\clearpage  % moving on to basal sliding

In our simulations, basal sliding is governed by a~pseudo-plastic sliding law,
already given in the manuscript but recalled here for the sake of completeness,
%
\begin{equation}
    \label{eqn:pseudoplastic}
    \vec{\tau}_b = -\tau_c \frac{\vec{v}_b}{{v_{th}}^q\,|\vec{v}_b|^{1-q}} \,,
\end{equation}
%
which relates the bed-parallel shear stresses, $\vec{\tau}_b$, to the sliding
velocity, $\vec{v}_b$. The yield stress, $\tau_c$, is modelled using the
Mohr--Coulomb criterion,
%
\begin{equation}
   \tau_c = c_0 + N\,\tan{\phi} \,,
\end{equation}
%
where cohesion, $c_0$, is assumed to be zero. Effective pressure, $N$, is
related to the overburden pressure, $P_0=\rho gh$, and the modelled amount of
subglacial water, using a formula derived from laboratory experiments with till
extracted from the base of Ice Stream B in West Antarctica
\citep[Eqn.~23]{Tulaczyk.etal.2000, Bueler.Pelt.2015},
\begin{equation}
    N = \delta P_0 \, 10^{(e_0/C_c) (1 - (W/W_{max}))} \,,
\end{equation}
where $\delta$ sets the minimum ratio between the effective and
overburden pressures, $e_0$ is a measured reference void ratio
and $C_c$ is a measured compressibility coefficient. The amount of water at the
base, $W$, varies from zero to $W_{max}$, a threshold above which additional
melt water is assumed to drain off instantaneously.

@@

In our sensitivity test, we set constant the pseudo-plastic sliding exponent,
$q=0.25$, and the threshold velocity, $v_{th}=100$\,\unit{m\,s^{-1}}, according
to values used by \citet{Aschwanden.etal.2013}, the till cohesion, $c_0=0$,
whose measured values are consistently negligible
\citep[p.~268]{Tulaczyk.etal.2000, Cuffey.Paterson.2010}, the till reference
void ratio, $e_0=0.69$, and the till compressibility coefficient, $C_c=0.12$,
according to the only measurements available to our knowledge by
\citep{Tulaczyk.etal.2000}.

We also use a constant spatial distribution of the till friction angle, $\phi$,
whose values vary from 15 to 45{\degree} as a piecewise-linear function of
modern bed elevation, with the lowest value occurring below modern sea level
(0\,m above sea level, m~a.s.l.) and the highest value occurring above the
generalised elevation of the highest shorelines
\citep[200\,m~a.s.l.,][Fig.~5]{Clague.1981}. This range of value span over the
range of measured values for glacial till of 18 to 40{\degree}
\citep[p.~268]{Cuffey.Paterson.2010}. It accounts for frictional basal
conditions associated with discontinuous till cover at high elevations, and
a~weakening of till associated with the presence of marine sediments at low
elevations (cf. \citealp{Martin.etal.2011};
\citealp[supplement]{Aschwanden.etal.2013}; \citealp{PISM-authors.2015}).

An additional simulation with a constant till friction angle, $\phi=30\degree$,
corresponding to the average value in \citet[p.~268]{Cuffey.Paterson.2010}, was
actually performed, but the induced variability was small as compared to that
which will be presented here, and therefore we decided to not include this run.

On the other hand, we test different values for the minimum ratio between the
effective and overburden pressures, $\delta$, and the maximum water
height in the till, $W_{max}$, as follow.

### 2015-12-30 reviews/response-jarosch.tex

\begin{table*}
  \caption{Parameter values used in the sensitivity test.}
  \label{tab:sens_params}
  \centering\makebox[\textwidth]
  {\begin{tabular}{l|ccccc|cc|cc}
    \tophline
            & \multicolumn{5}{c|}{Rheology}
            & \multicolumn{2}{c|}{Sliding}
            & \multicolumn{2}{c}{GRIP scaling} \\
    Config. & $A_c$ & $A_w$ & $Q_c$ & $Q_w$ & $E_{SIA}$
            & $\delta$ & $W_{max}$ & $f$ & $T_{[32, 22]}$ \\
            & \multicolumn{2}{c}{(\unit{Pa^{-3}\,s^{-1}})}
            & \multicolumn{2}{c}{(\unit{J\,mol^{-1}})}
            & & & (m) \\
    \middlehline
    Default$^1$  & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.02 & 2 & 0.38 & 6.2 \\
    \middlehline
    Soft ice$^2$ & $2.847\times 10^{-13}$
                 & $2.356\times 10^{-2}$
                 & $   60\times 10^3$
                 & $  115\times 10^3$
                 & 5 & 0.02 & 2 & 0.40 & 6.6 \\
    Hard ice$^2$ & $2.847\times 10^{-13}$
                 & $2.356\times 10^{-2}$
                 & $   60\times 10^3$
                 & $  115\times 10^3$
                 & 1 & 0.02 & 2 & 0.37 & 6.0 \\
    \middlehline
    Soft bed     & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.01 & 1 & 0.40 & 6.5 \\
    Hard bed     & $ 3.61\times 10^{-13}$
                 & $ 1.73\times 10^3$
                 & $   60\times 10^3$
                 & $  139\times 10^3$
                 & 1 & 0.05 & 5 & 0.36 & 5.9 \\
    \bottomhline
  \end{tabular}}
  \belowtable{After $^1$\citet{Paterson.Budd.1982,Bueler.Pelt.2015};
              and $^2$\citet{Cuffey.Paterson.2010}.}
\end{table*}

\begin{figure}
    \centering
    \includegraphics{sens_plot_rheo}
    \caption{Ice softness parameter, $A$, as a function of pressure-adjusted
             temperature, $T_{pa}$, for the default \citet{Paterson.Budd.1982},
             hard ice \citet{Cuffey.Paterson.2010}, and soft ice
             \citet[with $E_{SIA}=5$]{Cuffey.Paterson.2010} rheologies, using
             a linear scale (top panel) and logarithmic scale (bottom panel).
             Figure made using Eqn.~\ref{eqn:softness} with parameters from
             Table~\ref{tab:sens_params}.}
    \label{fig:sens_plot_rheo}
\end{figure}

@@

\clearpage  % moving on to basal sliding

In our simulations, basal sliding is governed by a~pseudo-plastic sliding law,
already given in the manuscript but recalled here for the sake of completeness,
%
\begin{equation}
    \label{eqn:pseudoplastic}
    \vec{\tau}_b = -\tau_c \frac{\vec{v}_b}{{v_{th}}^q\,|\vec{v}_b|^{1-q}} \,,
\end{equation}
%
which relates the bed-parallel shear stresses, $\vec{\tau}_b$, to the sliding
velocity, $\vec{v}_b$. The yield stress, $\tau_c$, is modelled using the
Mohr--Coulomb criterion,
%
\begin{equation}
   \tau_c = c_0 + N\,\tan{\phi} \,,
\end{equation}
%
where cohesion, $c_0$, is assumed to be zero. Effective pressure, $N$, is
related to the overburden pressure, $P_0=\rho gh$, and the modelled amount of
subglacial water, using a formula derived from laboratory experiments with till
extracted from the base of Ice Stream B in West Antarctica
\citep[Eqn.~23]{Tulaczyk.etal.2000, Bueler.Pelt.2015},
\begin{equation}
    \label{eqn:ntil}
    N = \delta P_0 \, 10^{(e_0/C_c) (1 - (W/W_{max}))} \,,
\end{equation}
where $\delta$ sets the minimum ratio between the effective and
overburden pressures, $e_0$ is a measured reference void ratio
and $C_c$ is a measured compressibility coefficient. The amount of water at the
base, $W$, varies from zero to $W_{max}$, a threshold above which additional
melt water is assumed to drain off instantaneously.

@@

\begin{itemize}
    \item{Our \emph{default} configuration used in the control run and all other
          simulations in the manuscript include $\delta=0.02$ and
          $W_{max}=2$\,m as in \citet{Bueler.Pelt.2015}.}
    \item{Our \emph{soft bed} configuration use $\delta=0.01$ and
          $W_{max}=1$\,m.}
    \item{Our \emph{hard bed} configuration use $\delta=0.05$ and
          $W_{max}=5$\,m.}
\end{itemize}

The effect of the three different parametrisations on the effective pressure on
the till, $N$, in response to water content, $W$, is illustrated in
Fig.~\ref{fig:sens_plot_ntil}.
All parameter choices are recalled in Table~\ref{tab:sens_params}.

Finally, we adjusted the GRIP linear scaling factor for each run, so that they
result in a similar glaciated area at the Last Glacial Maximum (LGM) to that
modelled with the default configuration (Table~\ref{tab:sens_params}).
In other words, the sensitivity in modelled ice sheet extent at MIS~2 is
expressed through the scaling factor required to obtain a fixed target area,
since we consider this scaling factor as a free parameter in our study.

\begin{figure}
    \centering
    \includegraphics{sens_plot_ntil}
    \caption{Effective pressure, $N$, as a function of water height at the
             base, $W$, for the default ($\delta=0.02$, $W_{max}=2$\,m),
             hard bed ($\delta=0.05$, $W_{max}=5$\,m), and soft bed
             ($\delta=0.01$, $W_{max}=1$\,m) sliding parametrisations, using a
             linear scale (top panel) and a logarithmic scale (bottom panel).
             Calculations are made for an ice thickness, $h$, of 1000\,m.
             Figure made using Eqn.~\ref{eqn:ntil} with parameters from
             Table~\ref{tab:sens_params} after
             \citet[Fig.~1]{Bueler.Pelt.2015}.}
    \label{fig:sens_plot_ntil}
\end{figure}

### 2015-12-30 reviews/response-jarosch.tex

\clearpage  % eventually moving on to the next point

All the information detailed above, including
Eqns~\ref{eqn:glenslaw} and~\ref{eqn:softness}, default parameter values,
Tables~\ref{tab:sens_params} and~\ref{tab:sens_extrema}, and
Figs.~\ref{fig:sens_plot_rheo}--\ref{fig:sens_ts} will be included in Sect.~2
(Model setup) or in the new section (Sensitivity to ice flow parameters).
Relevant discussion points in Sect.~4(Comparison with the geologic record)
will be updated to account for these new results.

\referee{%
    Generally section~4 appears to be quite long and seems to re-summarize
    known geological evidence for the region. At times the language is quite
    speculative, for example P4162 L1 and 7, P4164 L18, P4171 L7, L9, L17, L19
    and L20 and so forth. I would recommend to shorten that section to focus
    only on the geological evidence which can or can not clearly be reproduced
    by the presented model and avoid extensive speculation on what the reasons
    for mismatch are, especially in the present form of the manuscript, where a
    sensitivity study of the model itself is completely missing. However I
    leave the choice of how much geological evidence is discussed in the
    manuscript entirely up to the authors.}

Our choice of language P4162 L1 and 7 (discussion of a the longitudinal shift
between the modelled and geological reconstructed eastern margin and, in turn,
the position of the main ice divide), P4164 L18 (discussion of the confined
modelled ice extent in agreement with the regional geomorphology), and P4171
L7, L9, L17, L19 and L20 (discussion of the modelled flow patterns and the
observed lineation pattern on Interior Plateau of British Columbia) is
intentional. So is our choice to discuss some aspects of the model results that
are not unequivocal, as a result of uncertain model parameters (as partly
addressed by our new sensitivity study) or, more often, due to processes not
included in the model.

Our rationale for including this ``extensive speculation'' takes into account
the fact that glacial geology is no exact science. In fact, it often has to be
speculative and to include a large part of interpretation in order to reach
conclusions that are often uncertain but without which the field could not move
forward. Because we expect that part of our readership will consist of glacial
geologists who have sometimes spent an entire scientific career into studying
some aspects of the Cordilleran ice sheet (cf. comment by A.~Stumpf in this
discussion), we chose to address some of the long-standing debates on
Cordilleran ice sheet glacial geology, such as the location of ice divides or
the enigmatic lineation pattern on the Interior Plateau.

This is not only to prevent over-interpretation of our model results by
researchers with little or no background in ice sheet modelling, but also to
hint at potential model improvements that may be needed to approach these
debates through numerical modelling. Thus we do not plan to shorten section~4.
